Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-09-09 ms.tex
Updates to the DM section

Table~\ref{table:Bmin} shows the results on the lower limit of the central magnetic field that range
for the VERITAS limits from $B_{0} = 0.5$ to $1.4\,\mu$G and for the {\em Fermi} limits from $B_{0}
= 1.1$ to $4.6\,\mu$G.\footnote{Note that a central magnetic field of $3\,\mu$G corresponds in the
  Coma cluster to a relative magnetic energy density of $X_B=0.005$.} Since these lower limits on
$B_0$ are below the values favored by Faraday RM for the entire parameter space spanned by
$\alpha_B$ and $\alpha$, the hadronic model is perfectly viable.  This indicates the power of future
gamma-ray observations of Coma in constraining parameters of the hadronic model.

@@

\section{Emission from Dark Matter Annihilations}
When the size of the integration region is increased, the limits on $\expval{\sigma v}$ result from a competition between the gain in the astrophysical factor $\expval{J}$ and the integrated background. For integration regions larger than 0.2$^{\circ}$ in radius, the astrophysical factor no longer compensate for the increased count of background events and the signal-to-noise deteriorates.

\section{Discussion and Conclusions}
VERITAS observed the Coma cluster of galaxies for a total of 18.6 hours of high-quality live time between March and May in 2008. [{\em add flux limits}]

>>>>

Table \ref{table:Bmin} shows the results on the lower limit of the central magnetic field that range for the VERITAS limits from $B_{0} = 0.5$ to $1.4\,\mu$G and for the {\em Fermi} limits from $B_{0} = 1.1$ to $4.6\,\mu$G.\footnote{Note that a central magnetic field of $3\,\mu$G corresponds in the Coma cluster to a relative magnetic energy density of $X_B=0.005$.} Since these lower limits on $B_0$ are below the values favored by Faraday RM for the entire parameter space spanned by $\alpha_B$ and $\alpha$, the hadronic model is perfectly viable.  This indicates the power of future gamma-ray observations of Coma in constraining parameters of the hadronic model.

@@

\section{Emission from Dark Matter Annihilations}
As already mentioned in the introduction, most of the mass in a galaxy cluster is in the form of DM. While the nature of this DM remains unknown, the most compelling theoretical candidate is a WIMP. The self-annihilations of WIMPs are expected to produce a flux of secondary gamma rays with a spectral signature that deviates significantly from the power-law spectrum observed in most conventional astrophysical sources. The spectrum will also feature a very sharp cutoff at the WIMP mass. Observations of these spectral features together with the expected difference in spatial distribution of the gamma-ray flux compared to conventional astrophysical sources allows for a unique means to indirectly detect DM.

The expected gamma-ray flux due to self-annihilations of WIMPs is given by 
\begin{equation}
\frac{d\Phi_{\gamma}(\Delta\Omega,E)}{dE}=\frac{\expval{\sigma v}}{8\pi m_{\chi}^{2}}\frac{dN(E,m_{\chi})}{dE} J(\Delta\Omega),
\label{eqn:WIMPflux}
\end{equation}
where $\expval{\sigma v}$ is the thermally averaged product of the total self-annihilation cross section and the WIMP velocity, $m_{\chi}$ is the WIMP mass, $dN(E,m_{\chi})/dE$ is the differential gamma-ray yield per annihilation, $\Delta\Omega$ is the observed solid angle, and $J$ is the astrophysical factor - a dimensionless factor which determines the DM annihilation rate and depends on the DM distribution.

Given the upper limit on the observed gamma-ray rate, $R_{\gamma}(95\%\ \mathrm{CL})$, we can place constraints on the WIMP parameter space $(m_{\chi}, \expval{\sigma v})$. From Eq. (\ref{eqn:WIMPflux}) we get
\begin{equation}
\expval{\sigma v}(95\%\ \mathrm{CL}) > R_{\gamma}(95\%\ \mathrm{CL}) \frac{8\pi m_{\chi}^{2}}{J(\Delta\Omega)}\frac{1}{\int^{m_{\chi}} dE A_{\mathrm{eff}}dN_{\gamma}/dE}
\end{equation}
The astrophysical factor $J$ can be expressed as
\begin{equation}
J(\Delta\Omega)=\left(\frac{1}{\rho_{c}^{2}R_{H}}\right) \int_{\Delta\Omega}d\Omega\int d\lambda \rho^{2}(\lambda,\Omega),
\end{equation}
where $\lambda$ represents the line-of-sight. Because the self-annihilation of a WIMP is a two-body process, $J(\Delta\Omega)$ is proportional to the square of the DM density $\rho$. Here, $J(\Delta\Omega)$ has been normalized to the square of the critical density, $\rho_{c}$, and the Hubble radius, $R_{H}$.

In this work we have adopted the DM density profile by Navarro-Frenk-White \citep[NFW;][]{article:NavarroFrenkWhite:1997},
\begin{equation}
\rho_{\chi}(r)=\rho_{s}\left(\frac{r}{r_{s}}\right)^{-1}\left(1+\frac{r}{r_{s}}\right)^{-2},
\end{equation} 
which parameterizes the distribution of DM by the scale radius $r_{s}$ and the scale density $\rho_{s}$. [how where these calculated?]

%When the size of the integration region is increased, the limits on $\expval{\sigma v}$ result from a competition between the gain in the astrophysical factor $\expval{J}$ and the integrated background. For integration regions larger than 0.2$^{\circ}$ in radius, the astrophysical factor no longer compensate for the increased count of background events and the signal-to-noise deteriorates.

\section{Discussion and Conclusions}
VERITAS observed the Coma cluster of galaxies for a total of 18.6 hours of high-quality live time between March and May in 2008. No significant excess of gamma-rays above a threshold of about 220 GeV was detected. Flux upper limits at 99\% confidence level were calculated for the cluster core and three member galaxies. The core was considered both as a point-like and spatialy extended source. The obtained flux upper limits were used to constrain properties of the cluster.
